---
title: "Factorial Design Case Study"
author: "Harshvardhan Bhosale"
date: "2024-02-04"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Factorial Design 

We are going to perfrom a case study on facorial design based on the HairEyeColor dataset available in R.This data set has 3 variables Gender, Hair coluur and the colour of individuals eye along with the number of individuals of the specific conditions that we'll consider as our yeild.

```{r cars}
datasets::Titanic
```

## Data Preparation.

The above data needs to be modified to perform a factorial design experiment.

```{r}
y1<-c(0,0,35,0,0,0,17,0)
type1<-rep('Child',8)
stat1<-rep('dead',8)
gend1<-c(rep('M',4),rep('F',4))
y2<-c(118,154,387,670,4,13,89,3)
type2<-rep('Adult',8)
stat2=stat1
y3<-c(5,11,13,0,1,13,14,0)
type3<-type1
stat3<-rep('survived',8)
y4<-c(57,14,75,192,140,80,76,20)
type4=type2
stat4=stat3

Age<-c(type1,type2,type3,type4)
Gender<-c(rep(gend1,4))
No.Indivividuals<-c(y1,y2,y3,y4)
Status<-c(stat1,stat2,stat3,stat4)

data<-data.frame(Age,Gender,Status,No.Indivividuals)
data
```

## Importing the necessary packages:
```{r}
library(performance)
library(car)
library(carData)
library(doebioresearch)
```

## Fitting the Model:
```{r}
attach(data)

model<-lm(No.Indivividuals~Age*Gender*Status) 
summary(model)
anova(model)
```

## Various test that are required:
# Shapiro Test for Normality
```{r}
shapiro.test(data$No.Indivividuals)  
```

# Levene's Test 
```{r}
leveneTest(No.Indivividuals~Age)
leveneTest(No.Indivividuals~Gender)
leveneTest(No.Indivividuals~Status)
```




