---
title: "Factor Anlaysis Case Study"
author: "Harshvardhan Bhosale"
date: "2024-02-06"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Preparing the data for factor analysis.

```{r cars}
data=mtcars
```

## Loading Packages
```{r}
library(psych)
```

## Principal Component Analysis 
We'll start by perfroming a principal component analysis on the data.

```{r}
cor.mat<-cor(data) # obtaining the correlation of the data 
bartlett.test(data) #null hypothesis rejected data eligible for factor anlaysis
pca1<-princomp(data,cor = T)
summary(pca1)
plot(pca1,type='l')
pca1$loadings
pca1$scores
```

As we can see from the scree plot that the first 3 components explain the most of the variation in the data.

## Eligibility of the data for a Factor Analysis
1) The data is adequately large and metric
2) The data contains multicollinearity as can be seen from the correlaton matrix
3) Bartlett's Test of Sphericity 
```{r}
bartlett.test(data) 
```
As we can see the null hypothesis has been rejected.
# Since the above 3 conditions are satisfied we can proceed for a factor analysis.


## Factor Analysis 
Now we'll perform a factor analysis on the data.
We'll generate 3 factors because like we saw in the PCA three components are enough to explain 90% of the variation.
```{r}
fact1<-principal(data,nfactors = 3,rotate = 'none') # no. of factors=3 from the scree elbow
summary(fact1)
fact1$loadings
fact1$uniquenesses #specific variance of the analysis 
fact1$communality #comunalties of the analysis 
a_1<-diag(fact1$uniquenesses)
l_1<-fact1$loadings 
b_1<-l_1 %*% t(l_1)
residual1<-cor.mat-b_1-a_1
```
Here we have used no rotation on the analysis thus we can see that the factor are difficult to interpret and not much sensible.

## MLE method 
Here we will use the Maxium Likelihood Method to perform a factor analysis.

```{r}
fact2<-factanal(data,factors = 3,rotation='none')
summary(fact2)
fact2$loadings
fact2$uniquenesses
communality<-1-fact2$uniquenesses
l_2<-fact2$loadings
a_2<-diag(fact2$uniquenesses)
b_2<-l_2%*%t(l_2)
residual2<-cor.mat-b_2-a_2
```
Similar to the previous method we have not used anly rotaions on the analysis.
Note the changes in the formula for communality.

## Applying various Rotation methods:
Here, we'll try various rotation methods available to us e.g. varimax, equamax, promax and quartimax.

# Varimax Rotation
```{r}
rot1<-principal(data,nfactors = 3,rotate = 'varimax') # varimax rotation
rot1$loadings
rot1$uniquenesses #specific variance of the analysis 
rot1$communality #comunalties of the analysis 
a_3<-diag(rot1$uniquenesses)
l_3<-rot1$loadings 
b_3<-l_3 %*% t(l_3)
residual3<-cor.mat-b_3-a_3
```

# Equamax Rotation
```{r}
rot2 <- principal(data, nfactors = 3, rotate = 'equamax') # equamax Rotation
rot2$loadings
rot2$uniquenesses #specific variance of the analysis 
rot2$communality #communalties of the analysis 
a_4<-diag(rot2$uniquenesses)
l_4<-rot2$loadings 
b_4<-l_4 %*% t(l_4)
residual4<-cor.mat-b_4-a_4
```

# Promax Rotation
```{r}
rot3 <- principal(data, nfactors = 3, rotate = 'promax') # equamax Rotation

rot3$loadings
```

#Quartimax Rotation
```{r}
rot4 <- principal(data, nfactors = 3, rotate = 'quartimax') # equamax Rotation

rot4$loadings
```

## Inference 
As we can see from the above rotations, we have taken 3 factors from the data mtcars.
In this case the Varimax rotation method has given us the best result in creating distinct datasets. In this case, since the dataset is for cars we can draw the following inferences
The factors have been formed by the form by which we generally segregate the cars in market e.g. economy cars(cars with good fuel efficiency), normal car and performance car(sports car).
Here, we can see that factor RC3 contains the type of variables that define performance vehicles since it has higher weightage to cyl, disp, hp and negative weightage to qsec, mpg, wt, etc. Similarly, RC2 defines economy vehicles since the weightage are opposite to the performance cars. 
And RC1 stands for normal cars.

