---
title: "Case study"
author: "Harshvardhan Bhosale Roll No. 945"
date: "2024-01-11"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r cars}
data=mtcars
head(data)
```

## Performing Principal Component Analysis

Now we are going to foollow various steps involved in performing a principal component analysis

## Covariance Matrix

```{r pressure, echo=FALSE}
cov.mat<-cov(data)
cov.mat
```

## Correlation Matrix
```{r}
cor.mat<-cor(data)   
cor.mat
```
## Getting Eigen-values basedon the correlation matrix
```{r}
eign<-eigen(cor.mat)
eig.val<-eign$values
eign
```
## Performing Principal Component Analysis 
This step provides us with the analysis and the summary of the analysis performed whcih gives us the standard deviation of each component and the proportional variances of each component 
```{r}
pca_result <- princomp(mtcars, cor = TRUE)
# Explore the PCA results
summary(pca_result)
```

## Obtaining the scores
```{r}
pca_result$scores
```

## Plotting the proportional Varaince of Components 
Here we will make a scree plot of the proportional variances of the components of the PCA which will help us in determining the significant components 
```{r}
# Scree plot to visualize the proportion of variance explained by each principal component
plot(pca_result, type = "l", main = "Scree Plot")
```
Here we can see an elbow near compnent 3 thus we can say that the components 1 , 2, and 3 are significant 

## Making a Biplot
Biplot helps us in identifying the clusters that are made by the individual obserations and helps us in identifying their influence in components 
```{r}
biplot(pca_result)
```

Thus here we can see that several types of different cars e.g. sports cars like Maserati Bora and Ferrari Dino form a group together and a similar pattern is followed by the rest of the cars 
